An Innovative Storage Stack Addressing Extreme Scale Platforms and Big Data Applications

### 2015-09-15 paper.tex
pared down to the PDSW limit

\documentclass[conference]{IEEEtran}
\pdfpagewidth=8.5in
\pdfpageheight=11in
%\usepackage{subfig}
\usepackage{subfigure}
\usepackage[pdftex]{}
\usepackage{graphicx}
%\usepackage{todonotes}
\usepackage{listings}
\lstset {% general command to set parameter(s)
         language=C,
     basicstyle=\footnotesize,               % print whole listing small
%     keywordstyle=\color{black}\bfseries, % underlined bold black keywords
%     identifierstyle =\color{black},  % nothing happens
%     commentstyle=\color{black}\emph, % white comments
     %stringstyle=\ttfamily,          % typewriter type for strings
%     stringstyle=\color{black},       % typewriter type for strings
         tabsize=4,
         showtabs=false,
     showstringspaces=false}
%can't figure this one out for particles bandwidth
%\usepackage[caption,label]{subfig}

@@

%\conferenceinfo{Cluster'14,} {September 17, 2014, Madrid, Spain}
%\CopyrightYear{2014}
%\crdata{978-1-4503-1103-8/11/11}
%\clubpenalty=10000
%\widowpenalty = 10000

@@

\author{
\IEEEauthorblockN{Jay Lofstead\IEEEauthorrefmark{1}, Ivo Jimenez\IEEEauthorrefmark{2}, Carlos Maltzahn\IEEEauthorrefmark{2}, Quincey Koziol\IEEEauthorrefmark{3}, John Bent\IEEEauthorrefmark{4}, Eric Barton\IEEEauthorrefmark{5}}
\IEEEauthorblockA{\IEEEauthorrefmark{1}Sandia National Laboratories gflofst@sandia.gov}
\IEEEauthorblockA{\IEEEauthorrefmark{2}University of California, Santa Cruz ivo@cs.ucsc.edu, carlosm@soe.ucsc.edu}
\IEEEauthorblockA{\IEEEauthorrefmark{3}HDF Group koziol@hdfgroup.org}
\IEEEauthorblockA{\IEEEauthorrefmark{4}EMC john.bent@emc.com}
\IEEEauthorblockA{\IEEEauthorrefmark{5}Intel eric.barton@intel.com}
}
\maketitle

\begin{abstract}
The DOE Extreme-Scale Technology Acceleration Fast Forward Storage and IO Stack
project is going to have significant impact on storage systems design within
and beyond the HPC community. With phase 1 of the project complete, it is an
excellent opportunity to explore the complete design and how it will address
the needs of extreme scale platforms.  This paper examines the overall design.
Each layer of the proposed stack is discussed in some detail along with
cross-cutting topics that impact multiple layers, such as transactions and
metadata management. We also describe the proposed features to better support
some classes of big data applications.

This paper not only provides a timely summary of important aspects of the
design specifications but also captures the underlying reasoning that is not
available elsewhere. With this information, we encourage the broader community
to understand the design, intent, and future directions to foster discussion
guiding phase 2 and the ultimate production storage stack based on this work.
An initial performance evaluation of the early prototype implementation is also
provided to validate the presented design.

@@

Current production HPC IO stack design is unlikely to offer sufficient features
and performance to adequately serve extreme scale science platform
requirements.  Adding to the problem complexity is the variety of Big Data
problems users want to address using these platforms. Trying to address both of
these domains is a daunting problem.

To address these challenges, a joint effort between the US Department of
Energy's Office of Advanced Simulation and Computing and Advanced Scientific
Computing Research commissioned a project to develop a design and prototype for
an IO stack suitable for the extreme scale environment. It will be referred to
as the Fast Forward Storage and IO (FFSIO) project. This is a joint effort led
by Lawrence Livermore National Laboratory, with the DOE Data Management Nexus
leads Rob Ross and Gary Grider as coordinators and contract lead Mark Gary. The
participating labs are LLNL, SNL, LANL, ORNL, PNL, LBNL, and ANL.  Additional
industrial partners contracted include the Intel Lustre team, EMC, DDN, and the
HDF Group. This team has developed a specification
set~\cite{fastforward:2014:docs} for a future IO stack to address the
identified challenges. The first phase recently completed with a second phase
currently getting underway as of early 2014. The core focus of the first phase
was basic functionality and design. While an idealized potential system would
be the perfect target architecture, the reality of budgets has tempered many of
the decisions. For example, extensive availability of NVRAN or SSDs on all of
the compute nodes is currently not economically feasible limiting some of the
potential design choices.  With this in mind, the second phase will refine this
design incorporating fault recovery and other features missing from the first
phase.

@@

Overall, the architecture shifts from the idea of files and directories to
containers of objects. This shift avoids the bottlenecks related to the POSIX
files and directories structure such as the serialization of file creation, the
limitation and impact of the number of files in a directory, and the limited
semantics of a byte stream. Instead, the new interface focuses on high-level
data models and their properties and relationships. This concept permeates the
entire IO stack.

In addition to addressing the traditional scientific workload, this project
seeks to expand functionality to better support Big Data type applications. The
key idea is to support Arbitrary Connected Graphs (ACGs) such as those used in
Map-Reduce systems. Key system features are introduced to efficiently support
these computing models in addition to the typically bursty IO loads of more
traditional HPC applications.

\begin{figure}[htbp]
\vspace{-0.10in}
\centering
\includegraphics[width=\columnwidth]{images/arch-mapping}
\vspace{-0.15in}
\caption{Target Architecture and Component Mapping}
\label{fig:arch-mapping}
\vspace{-0.15in}
\end{figure}

At a more detailed view, the various layers of the IO stack each contribute
different functionality.  The architecture (Figure~\ref{fig:arch-mapping})
incorporates five layers, some of which have potentially optional components.
The top layer is generally a high level IO library, such as the demonstration
HDF5 library~\cite{hdf5}. It is in dark blue. The intent is to only have
access to the storage stack through such an API to manage the complexity of
working with the lower layers and to enable advanced functionality that would
require more direct user intervention. This layer incorporates the necessary
features for ACGs from a end-user's perspective.

Below the user API is an IO forwarding layer that redirects IO calls from the
compute nodes to the IO dispatching layer (in black).  This IO forwarding layer
is analogous to the function of the IO nodes in a BlueGene machine or the
passive data staging processes demonstrated
previously~\cite{nisar:2008:staging,Abbasi:2009:datatap}. The next two layers
have considerable functionality. The IO dispatcher (IOD) serves as the primary
storage interface for the IO stack (in green) and offers features like Burst
Buffers to insulate the persistent storage array from bursty IO workloads.
Ideally, the IOD layer's functionality can be optional based on available
hardware and compute power provided on the IO Nodes (IONs). Much of the
functionality offered at this layer would shift either up or down the stack as
discussed in detail below. The Distributed Asynchronous Object Storage (DAOS)
layer serves as the persistent storage interface and translation layer between
the user-visible object model and the requirements of the underlying storage
infrastructure. It is intended to be the traditional file system-like
foundation on which everything else is built with no dependence on any
technologies specified above it (in dark pink and yellow). For example, the IOD
layer with or without burst buffers is not required for DAOS to operate
properly.  Instead, the DAOS layer can handle all of the IO operations from the
user API layer, albeit with the potential performance penalty of manipulating
the shared, persistent storage array. At the bottom is the Versioning Object
Storage Device (VOSD) (in purple).  It serves as the interface for storing
objects of all types efficiently for each storage device in the parallel
storage array. Think of this layer as the physical disk interface layer. In
terms of Lustre, this would replace the API on individual storage devices with
an interface friendlier to the containers of objects and transactions/epochs
concepts used in the higher layers.

Along with analysis of the published design documents, a discussion of the
design philosophy representing the overall intent is presented. This
information represents information that may or may not have been written down,
but is the intent of ultimate product.  These ideas are presented to give a
fuller picture of where the project is going rather than dwelling on any
limitations of the published documents. This is most important to illustrate
how different concepts will work across layers since that information is spread
across multiple documents and may lack a cohesive overall view.

The rest of the paper is organized as follows. A brief overview of related work
is presented first in Section~\ref{sec:related}. Section~\ref{sec:end-user}
discusses the programmatic interface end users will see when interacting with
the storage array. This will be discussed in the context of the HDF5 based
example library used for the functionality demonstration. Section~\ref{sec:iof}
briefly discusses the motivation and proposal for the IO forwarding layer.
Section~\ref{sec:iod} describes the IO Dispatcher layer and the broad
functionality it offers. This will detail the pieces of the layer that are
potentially optional and mention the cross-cutting features discussed in a
later, cross-cutting section. Section~\ref{sec:daos} discusses how the DAOS
layer functions. As with the IOD layer, the cross-cutting features will be
mentioned, but discussed more fully in the cross-cutting section. The VOSD
layer is discussed in Section~\ref{sec:vosd}. In particular, the mapping
between the DAOS and VOSD layers are explored as it pertains to the physical
storage. Next is an exploration of cross-cutting features like transactions and
metadata management in Section~\ref{sec:broader}. Since these and other
features are spread across multiple layers, it makes more sense to discuss them
independently once an understanding of the overall structure has been
presented.  A demonstration of the functionality is presented in
Section~\ref{sec:evaluation}. This shows that the prototype system based on the
proposed design can function. The paper is concluded in
Section~\ref{sec:conclusion} with a summary of the broad issues covered in the
paper.

@@

Many projects over the last couple of decades have sought to address some
challenging aspect of parallel file system design. The recent rise of ``Big
Data'' applications with different characteristic IO patterns have somewhat
complicated the picture. Extreme scale machines will be expected to handle both
the traditional simulation-related workloads as well as applications more
squarely in the Big Data arena. This will require some adjustments to the
underlying system for good performance for both scenarios.

The major previous work is really limited to full file systems rather than the
mountain of file system refinements made over the years. A selection of these
other file systems and some features that make it relatively unique are
described below.

Ceph~\cite{weil:ceph} is a distributed object store and file system. It offers
both a POSIX and object interface including features typically found in parallel
file systems. Ceph's unique striping approach uses pseudo-random numbers with a
known seed eliminating the need for the metadata service to track where each
stripe in a parallel file is placed.

PVFS~\cite{carns:pvfs} offers optimizations to reduce metadata server load,
such as a single process opening a file and sharing the handle. It has been
commercialized in recent years as OrangeFS.

Lustre~\cite{braam:lustre-arch} has become the de facto standard on most major
clusters offering scalable performance and fine-grained end-user and
programmatic control over how data is placed in the storage system.

GPFS~\cite{schmuck:gpfs} offers a hands-off approach for providing good
performance for scaling parallel IO tasks and is used extensively by its owner,
IBM.

Panasas~\cite{panasas:architecture} seeks to offer a dynamically adaptive
striping system that detects the need for additional stripes for performance
and adjusts the file layout as necessary.

Other file systems, like GoogleFS~\cite{ghemawat:googlefs} and
HDFS~\cite{Shvachko:2010:hdfs}, address distributed rather than parallel
computing and cannot be compared directly. The primary difference between
distributed and parallel file systems is the ability of the file system to
store and retrieve data simultaneously from multiple clients, in parallel, and
treat the resulting collection of pieces as a single object.  Distributed file
systems rely on a single client creating a file, but distributing the set of
files across a wide array of storage devices. The other, popular distributed
file system of note is NFS~\cite{powlowski:1994:nfs3} that has been used for
decades for enterprise file systems.  These other file systems are mainly of
interest in the context of the ACG features of FFSIO and will be discussed more
in Section~\ref{sec:acg}.

@@

Since the proposal specifies a high-level IO API will be the primary end-user
interface for programmatically interacting with the FFSIO stack, the team used
the HDF5 API and leveraged the Virtual Object Layer (VOL) for the initial
design and implementation demonstration. This also serves as a good test
determining what are strictly necessary extensions to an existing IO API to
support the new functionality.  The additional functionality, such as
transactions, can be ignored for legacy implementations, but these applications
will not be able to take advantage of the asynchronous IO support inherent to
the new API.  The additions comprise (Figure~\ref{fig:vol-arch}):

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt

\item
  API extensions to support new functionality provided by the FFSIO project.
  This includes calls for managing asynchronous request lists, performing
  asynchronous operations, creating and managing transactions, end-to-end data
  integrity, and data type and functions to support the ACG functionality more
  efficiently than the current API.

\item\label{item:fs}
  Function shipping from Compute Nodes (CN) to IO Nodes (ION). This provides
  the application developer with the capability of sending computation down to
  the IONs and get back results and perform other operations such as indexing
  and data reorganization for more efficient retrieval.

\item
  Analysis Shipping from CN to IONs or DAOS nodes. This is similar
  to~\ref{item:fs} but instead of returning the result over the network, it
  gets stored on the nodes and pointers to it are returned.

\end{enumerate}

Function and Analysis Shipping are part of the cross-cutting features and are
discussed in Section~\ref{sec:fn-shipping}.

HDF5~\cite{hdf5} has a versatile data model offering complex data objects
and metadata. Its information set is a collection of datasets, groups,
datatypes and metadata objects. The data model defines mechanisms for
creating associations between various information items. The main
components of HDF5 are described below.

\begin{itemize}

\item
  \textbf{File}: In the HDF5 data model, the collection of data items stored
  together is represented by a file. It is an object collection that also
  describes the relationship between them. Every file begins with a root
  group ``/'' serving as the ``starting-point'' in the object hierarchy.

\item
  \textbf{Group}: A group is an object allowing association between HDF5
  objects. It is synonymous with directories in a file system. A group
  could contain multiple other groups, datasets, datatypes or attributes within
  it.

\item
  \textbf{Dataset}: HDF5 datasets are objects representing actual data
  or content. Datasets are arrays with potentially multiple dimensions. A
  dataset is characterized by a dataspace and a datatype. The dataspace
  captures the rank (number of dimensions) and the current and maximum
  extent in each dimension. The datatype describes the type of its data
  elements.

\item
  \textbf{Attribute}: Attributes are used for annotating HDF5 objects. They are
  datasets themselves and are attached to existing objects.

\end{itemize}

\subsection{Virtual Object Layer}
\label{virtual-object-layer}

The Virtual Object Layer is an abstraction mechanism internal to the HDF5
library~\cite{hdf5}. As shown in Figure~\ref{fig:vol-arch} it is implemented
just below the public API. The VOL exports an interface that allows writing
plugins for HDF5 enabling developers to handle data in ways other than writing
to storage in an HDF5 format.  Plugin writers provide an implementation for a
set of functions and are trusted to provide the proper semantics for the new
environment. For example, data staging could be implemented in the VOL layer by
replacing writing to disk in the HDF5 format to sending data to a data staging
area using some messaging mechanism.

For this project, rather than the default writing to disk in the HDF5 format,
the VOL is used to interact with the IOD layer and the different concepts it
offers without requiring all of the functionality be exposed to users. For
example, the containers and objects concept is transparently mapped to the
files and datasets existing HDF5 users are familiar with.  This reduces the
difficulty porting applications to the new IO stack.

\begin{figure}[htbp]
\vspace{-0.10in}
\centering
\includegraphics[width=\columnwidth]{images/vol-arch.png}
\vspace{-0.20in}
\caption{Architectural view of the VOL abstraction mechanism}
\label{fig:vol-arch}
\vspace{-0.10in}
\end{figure}

The IOD VOL plugin serves as the bridge between HDF5 and the IOD or DAOS Layer
(Figure~\ref{fig:vol-arch}). The application calls the HDF5 library while
running on the system's compute nodes. Using the VOL architecture, the IOD VOL
plugin uses a function shipper (RPC library) to forward the VOL calls to a
server component running on the IO nodes (IONs). This function shipping is the
IO Forwarding Layer discussed briefly in Section~\ref{sec:iof}. Once the calls
arrive at the IO nodes, they are translated into IO Dispatcher (IOD) API calls
and executed at the IONs.

@@

Since HDF5 has traditionally offered an interface focused on files and the
internal data types, such as datasets, these concepts must be mapped onto the
proposed FFSIO data storage concepts. At a high level, think of the HDF5 types
mapping to FFSIO types as follows: file $\rightarrow$ container, dataset $\rightarrow$
array, and groups and attributes $\rightarrow$ key-value store.
%This mapping is shown in Table~\ref{tab:mapping}

%\begin{table}[ht]
%    \vspace{-0.10in}
%    \centering
%    \caption[HDF5 to FFSIO Mapping]{HDF5 Data Model to FFSIO Data Model Mapping}
%    \bigskip
%    \vspace{-0.15in}
%
%    \begin{tabular}{|r|r|}
%\hline
%HDF5 & FFSIO\\
%\hline
%file & container \\
%dataset & array \\
%group & key-value store \\
%attribute(s) & key-value store \\
%\hline
%    \end{tabular}
%    \label{tab:mapping}
%\end{table}
%
In Section~\ref{sec:iod} below, the FFSIO types are described in more detail.

@@

The IO Forwarding layer offers a mechanism to reduce the concurrency impact of
the massive process count on the storage stack. A current trend mixing MPI with
a threading library like OpenMP is addressing the same issue, but limited to
handling the parallelism on a single node rather than multiple nodes. Projected
extreme scale platforms will have far fewer storage stack end-points per
compute process in which to receive requests and data. By reducing the number
of simultaneous requests, delays can be reduced. This has been demonstrated for
the file open operation with Lustre~\cite{lofstead:2009:adaptable} and to some
degree for accessing the storage devices
themselves~\cite{lofstead:2010:io-variability}.  The BlueGene platform
incorporated dedicated hardware to perform this role. The proposed
functionality for this layer, beyond managing the number of connections to the
IOD layer, is to implement function shipping from the compute nodes to the IO
nodes.

@@

\subsubsection{Container}
As mentioned above, the concept of a container is similar to that of a file in
a traditional file system. However, rather than being in a directory structure,
each container essentially is stored in a hash-space allowing direct access to
any container without regard to the current organizational context of the
file system. For example, there is no need to navigate a directory hierarchy
to name a particular container.

>>>>

\documentclass{sig-alt-gov2}
\pdfpagewidth=8.5in
\pdfpageheight=11in
%\usepackage{subfig}
\usepackage{subfigure}
\usepackage[pdftex]{}
\usepackage{graphicx}
%\usepackage{todonotes}
\usepackage{listings}
\lstset {% general command to set parameter(s)
         language=C,
     basicstyle=\footnotesize,               % print whole listing small
%     keywordstyle=\color{black}\bfseries, % underlined bold black keywords
%     identifierstyle =\color{black},  % nothing happens
%     commentstyle=\color{black}\emph, % white comments
     %stringstyle=\ttfamily,          % typewriter type for strings
%     stringstyle=\color{black},       % typewriter type for strings
         tabsize=4,
         showtabs=false,
     showstringspaces=false}
%can't figure this one out for particles bandwidth
%\usepackage[caption,label]{subfig}

@@

%\conferenceinfo{PDSW'15,} {November 17, 2015, Austin, Texas}
%\CopyrightYear{2015}
%\crdata{978-1-4503-1103-8/11/11}
%\clubpenalty=10000
%\widowpenalty = 10000

@@

\numberofauthors{1}
\author{
\alignauthor {Jay Lofstead\footnotemark[1], Ivo Jimenez\footnotemark[2], Carlos Maltzahn\footnotemark[2], Quincey Koziol\footnotemark[3], John Bent\footnotemark[4], Eric Barton\footnotemark[5]}\\
\footnotemark[1]~~Sandia National Laboratories gflofst@sandia.gov\\
\footnotemark[2]~~University of California, Santa Cruz ivo@cs.ucsc.edu, carlosm@soe.ucsc.edu\\
\footnotemark[3]~~HDF Group koziol@hdfgroup.org\\
\footnotemark[4]~~EMC john.bent@emc.com\\
\footnotemark[5]~~Intel eric.barton@intel.com
}
\maketitle

\begin{abstract}
The DOE Extreme-Scale Technology Acceleration Fast Forward Storage and IO Stack
project will significantly impact storage systems design within and beyond the
HPC community. With phase 2 starting, it is an excellent opportunity to explore
the complete design and how it will address the needs of extreme scale
platforms.
%Each layer of the proposed stack is discussed in some detail along
%with cross-cutting topics that impact multiple layers, such as transactions and
%metadata management.
%We also describe the proposed features to better support
%some classes of big data applications.
Key underlying reasoning driving design
decisions not available elsewhere are presented.  We encourage the broader
community to understand the design, intent, and future directions to foster
discussion guiding phase 2 and the ultimate production storage stack based on
this work.  An initial performance evaluation of the prototype implementation
validates the presented design.

@@

%Current production HPC IO stack design is unlikely to offer sufficient features
%and performance to adequately serve extreme scale science platform
%requirements.  Adding to the problem complexity is the variety of Big Data
%problems users want to address using these platforms. Trying to address both of
%these domains is a daunting problem.

%To address these challenges,
A joint effort between the US Department of
Energy's Office of Advanced Simulation and Computing and Advanced Scientific
Computing Research commissioned a project to develop a design and prototype for
an IO stack suitable for the extreme scale environment. It will be referred to
as the Fast Forward Storage and IO (FFSIO) project. This is a joint effort led
by Lawrence Livermore National Laboratory, with the DOE Data Management Nexus
leads Rob Ross and Gary Grider as coordinators and contract lead Mark Gary. The
participating labs are LLNL, SNL, LANL, ORNL, PNL, LBNL, and ANL.  Additional
industrial partners contracted include the Intel Lustre team, EMC, DDN, and the
HDF Group. This team has developed a specification
set~\cite{fastforward:2014:docs} for a future IO stack to address the
identified challenges. The core focus of the first phase
was basic functionality and design. While an idealized potential system would
be the perfect target architecture, the reality of budgets has tempered many of
the decisions.
%For example, extensive availability of NVRAM or SSDs on all of
%the compute nodes is currently not economically feasible limiting some of the
%potential design choices.  With this in mind, the second phase will refine this
%design incorporating fault recovery and other features missing from the first
%phase.

@@

%Overall, the architecture shifts from the idea of files and directories to
%containers of objects. This shift avoids the bottlenecks related to the POSIX
%files and directories structure such as file creation serialization, the
%file count in a directory limitation and impact, and the limited
%semantics of a byte stream. Instead, the new interface focuses on high-level
%data models and their properties and relationships. This concept permeates the
%entire IO stack.

%In addition to addressing the traditional scientific workload, this project
%seeks to expand functionality to better support Big Data type applications. The
%key idea is to support Arbitrary Connected Graphs (ACGs) such as those used in
%Map-Reduce systems. Key system features are introduced to efficiently support
%these computing models in addition to the typically bursty IO loads of more
%traditional HPC applications. These features are not discussed for space
%reasons.

\begin{figure}[htbp]
\vspace{-0.10in}
\centering
\includegraphics[width=\columnwidth]{images/arch-mapping}
\vspace{-0.20in}
\caption{Architecture and Component Mapping}
\label{fig:arch-mapping}
\vspace{-0.15in}
\end{figure}

At a more detailed view, the various layers of the IO stack each contribute
different functionality.  The architecture (Figure~\ref{fig:arch-mapping})
incorporates five layers, some of which have potentially optional components.
The top layer is generally a high level IO library, such as the demonstration
HDF5 library~\cite{hdf5}. It is in dark blue. The intent is to only have
access to the storage stack through such an API to manage the complexity of
working with the lower layers and to enable advanced functionality that would
require more direct user intervention.
%This layer incorporates the necessary
%features for ACGs from a end-user's perspective.

Below the user API is an IO forwarding layer that redirects IO calls from the
compute nodes to the IO dispatching layer (in black).  This IO forwarding layer
is analogous to the function of the IO nodes in a BlueGene machine or the
passive data staging processes demonstrated
previously~\cite{nisar:2008:staging,Abbasi:2009:datatap}. The next two layers
have considerable functionality. The IO dispatcher (IOD) serves as the primary
storage interface for the IO stack (in green) and offers features like Burst
Buffers to insulate the persistent storage array from bursty IO workloads.
%Ideally, the IOD layer's functionality can be optional based on available
%hardware and compute power provided on the IO Nodes (IONs). Much of the
%functionality offered at this layer would shift either up or down the stack as
%discussed in detail below.
The Distributed Asynchronous Object Storage (DAOS)
layer serves as the persistent storage interface and translation layer between
the user-visible object model and the requirements of the underlying storage
infrastructure. It is intended to be the traditional file system-like
foundation on which everything else is built with no dependence on any
technologies specified above it (in dark pink and yellow).
%For example, the IOD
%layer with or without burst buffers is not required for DAOS to operate
%properly.  Instead, the DAOS layer can handle all of the IO operations from the
%user API layer, albeit with the potential performance penalty of manipulating
%the shared, persistent storage array.
At the bottom is the Versioning Object
Storage Device (VOSD) (in purple).  It serves as the interface for storing
objects of all types efficiently for each storage device in the parallel
storage array.
%Think of this layer as the physical disk interface layer. In
%terms of Lustre, this would replace the API on individual storage devices with
%an interface friendlier to the containers of objects and transactions/epochs
%concepts used in the higher layers.

%Along with analysis of the published design documents, a discussion of the
%design philosophy representing the overall intent is presented. This
%information represents information that may or may not have been written down,
%but is the intent of ultimate product.  These ideas are presented to give a
%fuller picture of where the project is going rather than dwelling on any
%limitations of the published documents. This is most important to illustrate
%how different concepts will work across layers since that information is spread
%across multiple documents and may lack a cohesive overall view.

The rest of the paper is organized as follows. A brief overview of related work
is presented first in Section~\ref{sec:related}. Section~\ref{sec:end-user}
discusses the programmatic interface end users will see when interacting with
the storage array.
%This will be discussed in the context of the HDF5 based
%example library used for the functionality demonstration.
Section~\ref{sec:iof}
briefly discusses the motivation and proposal for the IO forwarding layer.
Section~\ref{sec:iod} describes the IO Dispatcher layer and the broad
functionality it offers.
%This will detail the pieces of the layer that are
%potentially optional and mention the cross-cutting features discussed in a
%later, cross-cutting section.
Section~\ref{sec:daos} discusses how the DAOS
layer functions.
%As with the IOD layer, the cross-cutting features will be
%mentioned, but discussed more fully in the cross-cutting section.
The VOSD layer is discussed in Section~\ref{sec:vosd}.
%In particular, the mapping
%between the DAOS and VOSD layers are explored as it pertains to the physical
%storage.
Next is an exploration of cross-cutting features like transactions and
metadata management in Section~\ref{sec:broader}.
%Since these and other
%features are spread across multiple layers, it makes more sense to discuss them
%independently once an understanding of the overall structure has been
%presented.
A demonstration of the functionality is presented in
Section~\ref{sec:evaluation}.
%This shows that the prototype system based on the
%proposed design can function.
The paper is concluded in
Section~\ref{sec:conclusion} with a summary of the broad issues covered in the
paper.

@@

%Many projects over the last couple of decades have sought to address some
%challenging aspect of parallel file system design. The recent rise of ``Big
%Data'' applications with different characteristic IO patterns have somewhat
%complicated the picture. Extreme scale machines will be expected to handle both
%the traditional simulation-related workloads as well as applications more
%squarely in the Big Data arena. This will require some adjustments to the
%underlying system for good performance for both scenarios.
%
%The major previous work is really limited to full file systems rather than the
%mountain of file system refinements made over the years. A selection of these
%other file systems and some features that make it relatively unique are
%described below.

Ceph~\cite{weil:ceph} is a distributed object store and file system. It offers
both a POSIX and object interface including features typically found in parallel
file systems.
%Ceph's unique striping approach uses pseudo-random numbers with a
%known seed eliminating the need for the metadata service to track where each
%stripe in a parallel file is placed.
PVFS~\cite{carns:pvfs} offers optimizations to reduce metadata server load,
such as a single process opening a file and sharing the handle.
%It has been
%commercialized in recent years as OrangeFS.
Lustre~\cite{braam:lustre-arch} has become the de facto standard on most major
clusters offering scalable performance and fine-grained end-user and
programmatic control over how data is placed in the storage system.
GPFS~\cite{schmuck:gpfs} offers a hands-off approach for providing good
performance for scaling parallel IO tasks and is used extensively by its owner,
IBM.
Panasas~\cite{panasas:architecture} seeks to offer a dynamically adaptive
striping system that detects the need for additional stripes for performance
and adjusts the file layout as necessary.

Other file systems, like GoogleFS~\cite{ghemawat:googlefs} and
HDFS~\cite{Shvachko:2010:hdfs}, address distributed rather than parallel
computing and cannot be compared directly.
%The primary difference between
%distributed and parallel file systems is the ability of the file system to
%store and retrieve data simultaneously from multiple clients, in parallel, and
%treat the resulting collection of pieces as a single object.  Distributed file
%systems rely on a single client creating a file, but distributing the set of
%files across a wide array of storage devices.
The other, popular distributed
file system of note is NFS~\cite{powlowski:1994:nfs3} that has been used for
decades for enterprise file systems.
%These other file systems are mainly of
%interest in the context of the ACG features of FFSIO and will be discussed more
%in Section~\ref{sec:acg}.

@@

Since the proposal specifies a high-level IO API will be the primary end-user
interface for programmatically interacting with the FFSIO stack, the team used
the HDF5 API and leveraged the Virtual Object Layer (VOL) for the initial
design and implementation demonstration. This also serves as a good test
determining what are strictly necessary extensions to an existing IO API to
support the new functionality.
%The additional functionality, such as
%transactions, can be ignored for legacy implementations, but these applications
%will not be able to take advantage of the asynchronous IO support inherent to
%the new API.
%The additions comprise (Figure~\ref{fig:vol-arch}):
%
%\begin{enumerate}
%\def\labelenumi{\arabic{enumi}.}
%\itemsep1pt\parskip0pt\parsep0pt
%
%\item
\textbf{API extensions} to support new functionality provided by the FFSIO project.
  This includes calls for managing asynchronous request lists, performing
  asynchronous operations, creating and managing transactions, end-to-end data
  integrity, and data type and functions to support the big data oriented functionality more
  efficiently than the current API.
%
%\item
\label{item:fs}
  \textbf{Function shipping} from Compute Nodes (CN) to IO Nodes (ION). This provides
  the application developer with the capability of sending computation down to
  the IONs and get back results and perform other operations such as indexing
  and data reorganization for more efficient retrieval.
%
%\item
  \textbf{Analysis Shipping} from CN to IONs or DAOS nodes. This is similar
  to~\ref{item:fs} but instead of returning the result over the network, it
  gets stored on the nodes and pointers to it are returned.

%\end{enumerate}

%Function and Analysis Shipping are part of the cross-cutting features and are
%discussed in Section~\ref{sec:fn-shipping}.

%HDF5~\cite{hdf5} has a versatile data model offering complex data objects
%and metadata. Its information set is a collection of datasets, groups,
%datatypes and metadata objects. The data model defines mechanisms for
%creating associations between various information items. The main
%components of HDF5 are described below.
%
%\begin{itemize}
%\itemsep1pt\parskip0pt\parsep0pt
%
%\item
%  \textbf{File}: In the HDF5 data model, the collection of data items stored
%  together is represented by a file. It is an object collection that also
%  describes the relationship between them. Every file begins with a root
%  group ``/'' serving as the ``starting-point'' in the object hierarchy.
%
%\item
%  \textbf{Group}: A group is an object allowing association between HDF5
%  objects. It is synonymous with directories in a file system. A group
%  could contain multiple other groups, datasets, datatypes or attributes within
%  it.
%
%\item
%  \textbf{Dataset}: HDF5 datasets are objects representing actual data
%  or content. Datasets are arrays with potentially multiple dimensions. A
%  dataset is characterized by a dataspace and a datatype. The dataspace
%  captures the rank (number of dimensions) and the current and maximum
%  extent in each dimension. The datatype describes the type of its data
%  elements.
%
%\item
%  \textbf{Attribute}: Attributes are used for annotating HDF5 objects. They are
%  datasets themselves and are attached to existing objects.
%
%\end{itemize}

%\subsection{Virtual Object Layer}
%\label{virtual-object-layer}

%As shown in Figure~\ref{fig:vol-arch} it is implemented
%just below the public API. The VOL exports an interface that allows writing
%plugins for HDF5 enabling developers to handle data in ways other than writing
%to storage in an HDF5 format.  Plugin writers provide an implementation for a
%set of functions and are trusted to provide the proper semantics for the new
%environment. For example, data staging could be implemented in the VOL layer by
%replacing writing to disk in the HDF5 format to sending data to a data staging
%area using some messaging mechanism.
%
%For this project, rather than the default writing to disk in the HDF5 format,
%the VOL is used to interact with the IOD layer and the different concepts it
%offers without requiring all of the functionality be exposed to users. For
%example, the containers and objects concept is transparently mapped to the
%files and datasets existing HDF5 users are familiar with.  This reduces the
%difficulty porting applications to the new IO stack.
%
%\begin{figure}[htbp]
%\vspace{-0.10in}
%\centering
%\includegraphics[width=\columnwidth]{images/vol-arch.png}
%\vspace{-0.20in}
%\caption{VOL Abstration Architecture}
%\label{fig:vol-arch}
%\vspace{-0.10in}
%\end{figure}
%
%The IOD VOL plugin serves as the bridge between HDF5 and the IOD or DAOS Layer
%(Figure~\ref{fig:vol-arch}). The application calls the HDF5 library while
%running on the system's compute nodes. Using the VOL architecture, the IOD VOL
%plugin uses a function shipper (RPC library) to forward the VOL calls to a
%server component running on the IO nodes (IONs). This function shipping is the
%IO Forwarding Layer discussed briefly in Section~\ref{sec:iof}. Once the calls
%arrive at the IO nodes, they are translated into IO Dispatcher (IOD) API calls
%and executed at the IONs.

@@

Since HDF5~\cite{hdf5} has traditionally offered an interface focused on files and the
internal data types, such as datasets, these concepts must be mapped onto the
proposed FFSIO data storage concepts. At a high level, think of the HDF5 types
mapping to FFSIO types as follows: file $\rightarrow$ container, dataset $\rightarrow$
array, and groups and attributes $\rightarrow$ key-value store.
Shifting from file-oriented to the FFSIO structure happens in a Virtual Object Layer (VOL) plug-in. The VOL is an abstraction mechanism internal to the HDF5
library that allows replacing the file IO operations with other calls.
%This mapping is shown in Table~\ref{tab:mapping}

%\begin{table}[ht]
%    \vspace{-0.10in}
%    \centering
%    \caption[HDF5 to FFSIO Mapping]{HDF5 Data Model to FFSIO Data Model Mapping}
%    \bigskip
%    \vspace{-0.15in}
%
%    \begin{tabular}{|r|r|}
%\hline
%HDF5 & FFSIO\\
%\hline
%file & container \\
%dataset & array \\
%group & key-value store \\
%attribute(s) & key-value store \\
%\hline
%    \end{tabular}
%    \label{tab:mapping}
%\end{table}
%
FFSIO types are discussed in Section~\ref{sec:iod}.

@@

The IO Forwarding layer offers a mechanism to reduce the concurrency impact of
the massive process count on the storage stack.
%A current trend mixing MPI with
%a threading library like OpenMP is addressing the same issue, but limited to
%handling the parallelism on a single node rather than multiple nodes. Projected
%extreme scale platforms will have far fewer storage stack end-points per
%compute process in which to receive requests and data. By reducing the number
%of simultaneous requests, delays can be reduced. This has been demonstrated for
%the file open operation with Lustre~\cite{lofstead:2009:adaptable} and to some
%degree for accessing the storage devices
%themselves~\cite{lofstead:2010:io-variability}.
The BlueGene platform
incorporated dedicated hardware to perform this role. The proposed
functionality for this layer, beyond managing the number of connections to the
IOD layer, is to implement function shipping from the compute nodes to the IO
nodes.

@@

\textbf{Container}
%As mentioned above, the concept of
A container is similar to that of a file in a traditional file system. However,
rather than being in a directory structure, each container essentially is
stored in a hash-space allowing direct access to
any container without regard to the current organizational context of the
file system.
%For example, there is no need to navigate a directory hierarchy
%to name a particular container.
%
%Functionally, a container plays the same role as a file in that it holds a
%collection of presumably related data intended to be accessed and manipulated
%as a unit.
Since this is extended from HDF5 files, the container could also be
viewed as a directory tree of objects where each directory entry specifies
either a sub-directory (group) or some data or attribute. For the IOD layer, it
is a collection of objects. For HDF5, interpreting the objects builds the
structure.
